%%
%% ACM RecSys 2026 — IDSS Merchant Agent Paper
%% Authors: Juli Huang, Hannah Clay, Thomas Sarda, Sajjad Beygi
%% Advisors: Prof. Negin Golrezaei, Prof. Amin Saberi
%% Stanford University, CS and MS&E
%%
\documentclass[sigconf]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

\setcopyright{acmlicensed}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{XXXXXXX.XXXXXXX}
\acmConference[RecSys '26]{ACM Conference on Recommender Systems}{September 28--October 2, 2026}{Minneapolis, Minnesota, USA}
\acmISBN{978-1-4503-XXXX-X/26/09}

\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}
\usepackage{algorithm}
\usepackage{algpseudocode}
\citestyle{acmauthoryear}

\begin{document}

\title[Reliable Conversational Recommendation for Agentic E-Commerce]{Reliable Conversational Recommendation in Agentic E-Commerce:\\
A Merchant Agent Framework for Limited-Catalog Shopping}

%% ---- BLIND REVIEW VERSION ----
%% Author information suppressed per double-blind review policy.
%%
%% CAMERA-READY TODO: uncomment the full author block below (through \renewcommand{\shortauthors}{Huang et al.})
%%   and DELETE the Anonymous Author block that follows it.
%% CAMERA-READY TODO: in \begin{acks} near the end, replace:
%%   "[details omitted for blind review]"
%%   with:
%%   "The authors thank Prof. Negin Golrezaei and Prof. Amin Saberi for advising
%%   and infrastructure support at Stanford University."
%%
%% \author{Juli Huang}
%% \affiliation{%
%%   \institution{Stanford University}
%%   \department{Department of Computer Science}
%%   \city{Stanford}\state{CA}\country{USA}}
%% \email{julih@stanford.edu}
%%
%% \author{Hannah Clay}
%% \affiliation{%
%%   \institution{Stanford University}
%%   \department{Department of Computer Science and Biocomputation Program}
%%   \city{Stanford}\state{CA}\country{USA}}
%% \email{hclay@stanford.edu}
%%
%% \author{Thomas Sarda}
%% \affiliation{%
%%   \institution{Stanford University}
%%   \department{Department of Management Science \& Engineering}
%%   \city{Stanford}\state{CA}\country{USA}}
%% \email{tsarda@stanford.edu}
%%
%% \author{Sajjad Beygi}
%% \affiliation{%
%%   \institution{Stanford University}
%%   \department{Department of Electrical Engineering}
%%   \city{Stanford}\state{CA}\country{USA}}
%% \email{beygi@stanford.edu}
%%
%% \renewcommand{\shortauthors}{Huang et al.}
\author{Anonymous Author(s)}
\affiliation{%
  \institution{Anonymous Institution}
  \country{Anonymous Country}}
\renewcommand{\shortauthors}{Anonymous et al.}

\begin{abstract}
Modern AI shopping agents increasingly operate over \emph{limited catalogs}---merchant-owned
product databases where web search is disabled and the system must recommend reliably
from a fixed set of products. This setting exposes a fundamental limitation of
raw LLM-based recommenders: they hallucinate product specifications, forget user
constraints across turns, and cannot guarantee that recommended products actually
exist in the merchant's inventory.

We present an Interactive Decision Support System (IDSS), a merchant agent
framework designed specifically for limited-catalog shopping. IDSS addresses
reliability through a hybrid architecture that routes \emph{correctness-critical}
operations---hard constraint filtering, brand exclusion, progressive relaxation---through
a relational database and a Neo4j knowledge graph (2,400 nodes; 8,500 edges),
while reserving LLM inference for natural language understanding and narrative
generation. A multi-turn preference elicitation agent maintains session-persistent
slot values, enforces brand exclusions across all turns, and detects mid-conversation
preference shifts. The system additionally unifies three emerging agentic commerce
protocols (MCP, UCP, ACP) behind a single backend, enabling any LLM shopping
agent to access merchant catalogs through standardized, reproducible tool calls---an
early instantiation of the agent-to-agent e-commerce paradigm.

The KG re-ranking overlay resolves complex queries (``best value,'' ``find similar'')
in 207~ms median, compared to 2,256--5,714~ms for equivalent LLM narrative
generation, an 11--28$\times$ speedup with deterministic, hallucination-free results.
Evaluated on a 43-query single-turn benchmark, IDSS achieves avg.\ score
$0.683 \pm 0.020$ (pass rate 90.7\%), outperforming the LLM-only baseline by
4.7\,pp and catalog-aware GPT-4o-mini ($0.662 \pm 0.032$) on quality while
achieving perfect catalog grounding (1.00 vs.\ 0.0 for web-augmented baselines).
On a 10-scenario multi-turn benchmark, IDSS ($0.604 \pm 0.013$) leads
catalog-bound GPT ($0.595 \pm 0.015$), Baseline~B ($0.554 \pm 0.005$), and
Gemini ($0.532 \pm 0.018$) across constraint accumulation, brand-exclusion
persistence, and preference-pivot scenarios.
\end{abstract}

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10002951.10003317.10003347.10003350</concept_id>
  <concept_desc>Information systems~Recommender systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10002951.10003317.10003371.10003386</concept_id>
  <concept_desc>Information systems~Question answering</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010147.10010257.10010293.10010294</concept_id>
  <concept_desc>Computing methodologies~Multi-agent systems</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10002951.10003317.10003347.10003360</concept_id>
  <concept_desc>Information systems~Social networks</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Recommender systems}
\ccsdesc[300]{Information systems~Question answering}
\ccsdesc[300]{Computing methodologies~Multi-agent systems}
\ccsdesc[100]{Information systems~Social networks}

\keywords{limited-catalog shopping, reliability, conversational recommendation,
agentic commerce, multi-agent systems, knowledge graph reasoning,
preference elicitation, constraint enforcement, Model Context Protocol,
Universal Commerce Protocol, semantic search}

\maketitle

%% ==========================================================================
\section{Introduction}
\label{sec:introduction}

Conversational recommender systems increasingly rely on large language models
(LLMs) to interact with users in natural language, elicit preferences, and
generate recommendations. A particularly important and underexplored deployment
setting is \emph{limited-catalog shopping}: a merchant agent operating over its
own fixed product database, with web search disabled, that must recommend
reliably from the merchant's actual inventory. This setting is economically
motivated---it reduces compute cost, gives merchants control over product data,
and avoids the latency and hallucination risks of web-augmented search---but
it places strict reliability demands on the recommender. LLMs, despite their
fluency, remain unreliable in this setting: they may hallucinate product
specifications not present in the catalog, forget earlier brand exclusions across
turns, or apply budget filters inconsistently. A recommender that returns
out-of-catalog products or violates a user's stated constraints may sound
plausible but fails the merchant's core reliability requirement.

This paper studies the following research question: \emph{How can a merchant
agent reliably recommend products from a fixed catalog while enforcing user
constraints across multi-turn interactions---without relying on web search or
hallucinated product data?}

To address this question, we propose a conversational recommender that combines
LLM-based interaction with deterministic candidate filtering, knowledge-graph
reasoning, and persistent tracking of user preferences across turns. The
central design principle is to assign each component the task it is best suited
for: the LLM handles natural-language understanding and response generation,
while correctness-critical operations---such as constraint enforcement,
candidate filtering, and cross-turn preference persistence---are delegated to
structured system components.

Empirically, this approach improves both recommendation reliability and overall
quality. On a single-turn benchmark, it outperforms a plain LLM baseline while
achieving perfect brand and filter compliance through SQL-based enforcement. On
a multi-turn benchmark, it achieves the strongest overall performance among the
compared methods and consistently preserves constraints under preference
accumulation, reversals, and budget changes. In addition, knowledge-graph
traversal substantially accelerates structural recommendation queries such as
substitutes and compatibility, while providing deterministic and auditable
reasoning paths.
%% [Verified accurate: Table~\ref{tab:latency} shows KG overlay at 207ms vs LLM narrative at 2,256--5,714ms (11--28x speedup, n=10 direct-call benchmark, April 2026).]



Figure~\ref{fig:system_overview} provides an overview of the proposed system
design. Our contributions are threefold.
\begin{enumerate}
    \item \textbf{Merchant agent design.} We present a hybrid merchant agent
    that combines LLM-based natural language understanding with deterministic
    constraint enforcement and knowledge-graph reasoning, enabling reliable
    product recommendation over a fixed, merchant-owned catalog without
    web search.

    \item \textbf{Multi-turn preference elicitation.} We develop a
    session-persistent preference elicitation mechanism that preserves user
    constraints across turns, enforces brand exclusions, and adaptively asks
    informative follow-up questions to refine recommendations across complex
    multi-turn interactions.

    \item \textbf{Evaluation framework.} We introduce a hybrid G-Eval
    framework that separates recommendation quality from constraint
    correctness, enabling faithful assessment of conversational recommenders
    in both single-turn and multi-turn limited-catalog settings.
\end{enumerate}

The proposed system is designed for emerging \emph{agent-to-agent e-commerce}
architectures, where every merchant deploys its own AI agent and shopping agents
interact with merchant agents directly via structured protocols---avoiding web
crawling and enabling low-latency, merchant-controlled information exchange.
To this end, IDSS unifies three competing agentic commerce protocols---the
Model Context Protocol (MCP, Anthropic), the Universal Commerce Protocol (UCP,
Google), and the Agent Commerce Protocol (ACP, OpenAI)---behind a single
backend~\cite{Anthropic2024MCP,OpenAI2025ACP,Google2026UCP}.
The core contribution of this paper, however, is the reliability framework
that makes limited-catalog recommendation trustworthy, not the protocol layer
itself.
%% System Overview Figure — two-box: Shopping Agent ↔ Merchant Agent
\begin{figure}[t]
\centering
\begin{tikzpicture}[
  box/.style={draw, rounded corners=3pt, minimum width=5.2cm, minimum height=0.48cm,
              font=\small, align=center, fill=white},
  groupbox/.style={draw, rounded corners=6pt, minimum width=5.6cm,
                   font=\small\bfseries, align=center},
  store/.style={draw, rounded corners=4pt, minimum width=1.65cm, minimum height=1.2cm,
                font=\scriptsize, align=center},
  arr/.style={-{Stealth[length=5pt]}, thick},
  node distance=0.26cm
]
%% --- User utterance ---
\node[box] (user) {User Utterance};

%% --- Shopping Agent frame ---
\node[groupbox, below=0.5cm of user, minimum height=3.2cm,
      fill=blue!6, label={[font=\small\bfseries]above:Shopping Agent}]
  (sabox) {};
\node[box, below=0.88cm of user, fill=blue!10]  (qr) {Query Rewriting};
\node[box, below=0.26cm of qr,   fill=blue!10]  (ir) {Intent Routing};
\node[box, below=0.26cm of ir,   fill=blue!10]  (pe) {Preference Elicitation \& Session State};
\node[box, below=0.26cm of pe,   fill=blue!10]  (rg) {Response Generation};

%% --- Protocol bridge ---
\node[font=\scriptsize\itshape, below=0.16cm of rg] (proto)
  {$\updownarrow$~MCP / UCP / ACP};

%% --- Merchant Agent frame ---
\node[groupbox, below=0.40cm of rg, minimum height=3.2cm,
      fill=orange!6,
      label={[font=\small\bfseries]above:Merchant Agent (limited catalog)}]
  (mabox) {};
\node[box, below=1.05cm of rg,  fill=orange!12] (sql) {SQL Hard Constraint Filtering};
\node[box, below=0.26cm of sql, fill=orange!12] (kg)  {Knowledge Graph Traversal (Neo4j)};
\node[box, below=0.26cm of kg,  fill=orange!12] (rel) {Progressive Relaxation};
\node[box, below=0.26cm of rel, fill=orange!12] (cc)  {Cart \& Checkout Operations};

%% --- Datastores ---
\node[store, below left=0.7cm and 1.0cm of cc,  fill=yellow!12] (pg) {
  \textbf{PostgreSQL}\\21k products\\SQL filters
};
\node[store, below=0.7cm of cc,                  fill=purple!10] (neo) {
  \textbf{Neo4j KG}\\2.4k nodes\\8.5k edges
};
\node[store, below right=0.7cm and 1.0cm of cc,  fill=red!8]    (redis) {
  \textbf{Redis}\\Cache\\Session
};

%% --- Arrows ---
\draw[arr] (user)  -- (qr);
\draw[arr] (qr)    -- (ir);
\draw[arr] (ir)    -- (pe);
\draw[arr] (pe)    -- (rg);
\draw[arr] (rg)    -- (proto);
\draw[arr] (proto) -- (sql);
\draw[arr] (sql)   -- (kg);
\draw[arr] (kg)    -- (rel);
\draw[arr] (rel)   -- (cc);
\draw[arr] (cc)    -- (pg);
\draw[arr] (cc)    -- (neo);
\draw[arr] (cc)    -- (redis);
\end{tikzpicture}
\caption{System overview: the \emph{shopping agent} (blue) handles natural-language
  understanding, preference elicitation, and response generation.
  The \emph{merchant agent} (orange) enforces hard constraints over the merchant's
  fixed catalog via SQL filtering, KG traversal, and progressive relaxation.
  All protocol communication (MCP / UCP / ACP) passes through the merchant agent.}
\label{fig:system_overview}
\end{figure}

%% ==========================================================================
\section{Related Work}
%% ==========================================================================

\paragraph{Conversational Recommender Systems.}
Conversational recommender systems (CRSs) elicit user preferences through
multi-turn interaction before producing recommendations~\cite{Jannach2021Survey}.
Early work framed CRS as an interactive preference-elicitation problem, using
bandit-based question selection and dialogue-driven recommendation
models~\cite{Christakopoulou2016,Sun2018}. More recent work integrates LLMs
into CRS pipelines for preference understanding, sub-task management, and
response generation~\cite{Friedman2023,Feng2023LLMCRS,Liu2023CRSLLM}.
Research has also emphasized retrieval augmentation, memory, and evaluation:
ReFICR augments LLM-based CRS with retrieval to improve item
grounding~\cite{Yang2024ReFICR}; MemoCRS studies memory-enhanced preference
modeling across sessions~\cite{Xi2024MemoCRS}; and recent evaluation work
argues that fluent responses alone are insufficient for assessing LLM-based
CRSs~\cite{Yang2024BehaviorAlignment}. Our work addresses a distinct failure
mode: reliably enforcing hard user constraints (e.g., excluded brands,
structured filters) while maintaining preference state across turns in
shopping-oriented conversations.

\paragraph{Knowledge Graphs and Structured Reasoning for Recommendation.}
Knowledge-aware recommenders use structured relations among items, attributes,
and entities to improve recommendation quality and explainability.
Representative methods include Collaborative Knowledge Base Embedding
(CKE)~\cite{Zhang2016CKE}, Knowledge Graph Convolutional Networks
(KGCN)~\cite{Wang2019KGCN}, and Knowledge Graph Attention Network
(KGAT)~\cite{Wang2019KGAT}, all of which incorporate graph structure into
learned item representations. Our use of a knowledge graph is primarily
operational: it supports explicit structural queries (substitutes,
compatibility), produces auditable reasoning paths, and avoids delegating
correctness-critical decisions to free-form LLM inference.

\paragraph{LLM-based Agents for Shopping and Web Interaction.}
A related line of work studies LLM agents acting in interactive environments,
including web navigation and shopping benchmarks. WebArena and AgentBench
evaluate whether agents can follow instructions and complete multi-step tasks
in realistic interactive settings~\cite{Zhou2023WebArena,Liu2023AgentBench}.
More recent benchmarks such as ShoppingBench evaluate end-to-end agent
performance in e-commerce environments with large product
catalogs~\cite{Wang2025ShoppingBench}. These works are complementary to ours:
they study the behavior of the \emph{shopping agent} operating over an external
interface, whereas we focus on the \emph{merchant-side recommender design}
that provides structured access to catalog, constraint, and compatibility
information.


%% ==========================================================================
\section{Problem Setting}
\label{sec:problem}

We study conversational recommendation in the \emph{limited-catalog setting}:
a merchant agent that operates over a fixed, merchant-owned product database
with web search disabled. The agent must recommend products exclusively from
this catalog, with the following formal constraints:

\begin{itemize}
  \item \textbf{Catalog grounding.} Every recommended product must exist in
  the merchant's database with a verified price and available inventory.
  \item \textbf{Hard constraint enforcement.} User-specified constraints
  (budget ceiling, excluded brands, minimum specifications) must be enforced
  exactly---not approximately---across all turns.
  \item \textbf{Multi-turn persistence.} Constraints specified in any previous
  turn must remain active until explicitly revoked by the user.
  \item \textbf{No hallucination.} Product attributes (RAM, GPU, price, brand)
  must be read from structured database records, not generated by the LLM.
\end{itemize}

This setting is motivated by the agent-to-agent e-commerce paradigm, where
each merchant deploys an AI agent over its own catalog, and shopping agents
interact with merchant agents via structured protocols rather than web search.
It is also practically important for direct-to-consumer and B2B merchants who
need deterministic, auditable recommendation behavior.

Users may specify both soft preferences and hard constraints through natural-language
interaction. Hard constraints include requirements such as budget limits,
excluded brands, and minimum product specifications. The recommender must
produce relevant recommendations while preserving these constraints across
multiple turns, even when the user refines, relaxes, or reverses earlier
preferences.

This setting introduces three interrelated challenges. First, user requirements
are often expressed incrementally rather than all at once. Second, some
requirements are correctness-critical and should not be violated. Third,
shopping queries frequently involve structural product relations such as
substitutes and compatibility, which are difficult to handle reliably through
text generation alone.

Our goal is to design a merchant agent that supports natural-language interaction
while maintaining persistent preference state, enforcing hard constraints
deterministically, and resolving structural queries through structured reasoning---
all within the merchant's fixed catalog.

\section{The Shopping Agent}
\label{sec:shopping-agent}

The shopping agent is the user-facing component responsible for natural
language understanding, preference elicitation, and conversation management.
It receives user utterances, rewrites them into structured canonical forms,
routes intents, extracts slot-value preferences, and generates grounded
conversational responses. All \emph{correctness-critical} operations---hard
constraint enforcement, brand exclusion, progressive relaxation, and
transactional actions---are delegated to the merchant agent
(\S\ref{sec:merchant-agent}), ensuring that actions with real downstream
consequences remain grounded in structured data rather than inferred
through free-form language generation.

\subsection{System Overview}

At each turn, the system receives the current user utterance together with the
dialogue history and returns either a conversational response, a product set,
or the result of an action such as comparison, add-to-cart, or checkout
progression. The end-to-end pipeline contains five stages: query rewriting,
intent routing, structured extraction, session-state update, and tool- or
catalog-grounded execution. This design allows the system to support a wide
range of shopping capabilities beyond recommendation alone, including product
search, comparison, refinement, cart operations, and transactional flows.

\subsection{Agentic LLM Framework}

The agentic framework is the entry point for each user turn. It acts as a
controller that interprets the user message in context and decides what kind of
operation should be performed next. Rather than issuing a single generative
response, the framework determines whether the user is requesting product
discovery, recommendation refinement, product comparison, similar-item
retrieval, add-to-cart, cart inspection, or checkout-related actions.

The first stage is \emph{query rewriting}. Before execution, the system rewrites
the user utterance into a more canonical form that exposes latent intent,
resolves ambiguity, and improves downstream extraction. This step supports
common-sense enrichment, contradiction detection, and adaptation to the user's
level of technical specificity.

The second stage is \emph{intent routing}. The system maps each turn to an
action type, such as new search, refinement, comparison, substitute request,
cart action, or transactional continuation. Some frequent intents are handled
through lightweight deterministic rules, while more ambiguous requests are
resolved by the LLM agent. This layered routing design reduces unnecessary LLM
calls for repetitive actions while preserving flexibility for open-ended
shopping dialogue.

\subsection{Structured Extraction and Session State}

After routing, the system extracts structured information relevant to the
current turn. Depending on the intent, this may include product preferences,
hard constraints, referenced items, comparison targets, or cart-related
entities. For recommendation-oriented turns, the extracted structure may
include budget, excluded brands, required features, and use-case indicators.
For action-oriented turns, it may include item identifiers, relative
references, quantities, or follow-up commands such as selecting a previously
shown product.

All extracted information is merged into a persistent session state. This state
stores the currently active constraints, the recent recommendation context, the
cart state, and previously referenced products. Persistent state is essential
for handling multi-turn shopping behavior, such as “show me something cheaper,”
“compare the first and third options,” or “add the second one to the cart.”
Without such state, these requests would be difficult to resolve reliably.

\subsection{Deterministic Commerce Layer}

The deterministic layer is responsible for executing correctness-critical
operations over structured data sources. This includes enforcing product
constraints, retrieving feasible products, resolving structural product
relations, maintaining cart contents, and supporting transactional flows. The
purpose of this layer is to ensure that actions with real downstream
consequences are grounded in explicit system state rather than inferred only
through free-form language generation.

For product retrieval, hard requirements such as budget limits, excluded
brands, product type, and minimum specifications are translated into structured
filters and applied directly at the catalog layer. For cart and checkout
operations, the deterministic layer resolves product references against the
current session and executes the requested action only after the target item is
identified unambiguously. This is particularly important for commands such as
add-to-cart, remove-from-cart, and checkout progression, where incorrect
resolution would directly affect user experience.

\subsection{Recommendation, Comparison, and Structured Reasoning}

Recommendation is one important capability of the system, but it is not the
only one. After deterministic filtering, the system ranks feasible candidates
using softer preference signals derived from use case, product attributes, and
embedding-based similarity. This supports recommendation and refinement tasks
such as finding a better fit, surfacing alternatives, or narrowing results
after follow-up preference updates.

The system also supports structured comparison and relation-based reasoning.
When the user asks to compare products, the system retrieves the referenced
items from session state or catalog results and assembles an attribute-grounded
comparison rather than relying on unconstrained LLM recall. Likewise, queries
such as “show me similar items,” “find a better option,” or “what accessories
work with this?” are resolved through knowledge-graph traversal over explicit
relations such as similarity, compatibility, manufacturer, and use case. This
provides a structured basis for both recommendation and comparison tasks.

\subsection{Multi-turn Interaction Management}

The system manages multi-turn interaction by deciding when to answer directly
and when to ask for clarification. If the user request is underspecified, the
agent asks a follow-up question designed to reduce uncertainty over the current
feasible set. For recommendation-oriented turns, question selection is based on
the expected informativeness of missing preference slots. When the user has
already supplied enough information, the system skips elicitation and proceeds
directly to execution.

This interaction policy also supports conversational continuity across action
types. A session may begin with broad exploration, move to refinement and
comparison, then continue to add-to-cart and checkout. The system therefore
treats the shopping flow as a connected decision process rather than as a
sequence of isolated recommendation queries.

\subsection{Response Generation and Action Realization}

The final stage generates a grounded conversational response. Depending on the
intent, the output may include recommended products, a comparison summary, a
clarification question, confirmation of a cart action, or the next step in a
checkout flow. Response generation is grounded in structured outputs from the
deterministic layer rather than relying on unconstrained LLM memory. As a
result, the LLM is used to communicate and organize results, while actual
product selection, action execution, and state transitions remain under
structured control.
\section{The Merchant Agent}
\label{sec:merchant-agent}

The merchant agent is the reliability layer of the system. It exposes the
merchant's fixed product catalog through structured protocol endpoints,
enforces hard constraints deterministically, and provides knowledge-graph-accelerated
resolution of complex product queries. All correctness-critical operations
execute here: SQL constraint filtering, brand exclusion, progressive
relaxation, and structured product-relation traversal.
By operating exclusively over the merchant's own catalog---with no web search---
the merchant agent guarantees that every recommended product exists in
inventory, has a known price, and satisfies the user's stated constraints.
It implements the agent-to-agent e-commerce paradigm by exposing MCP, UCP,
and ACP protocol families from a single backend
(\S\ref{sec:protocol-unification}), allowing any shopping agent to access
the catalog without web crawling.

Figure~\ref{fig:arch} shows the end-to-end architecture.
The system comprises two FastAPI services: a product data layer (\texttt{mcp-server},
port 8001) and an agent layer (\texttt{agent}, port 8000), backed by three
data stores.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  svc/.style={draw, rounded corners=4pt, minimum width=5.4cm,
              font=\small, align=center, fill=blue!8},
  store/.style={draw, rounded corners=4pt, minimum width=1.6cm, minimum height=1.4cm,
                font=\scriptsize, align=center},
  arr/.style={-{Stealth[length=4.5pt]}, thick},
  node distance=0.5cm
]
%% --- Agent Service ---
\node[svc, minimum height=1.2cm] (agent) {
  \textbf{Agent Service} \texttt{(:8000)}\\[2pt]
  \texttt{chat\_endpoint} $\cdot$ \texttt{universal\_agent}\\
  \texttt{query\_rewriter} $\cdot$ \texttt{comparison\_agent}
};

%% --- MCP Server ---
\node[svc, minimum height=1.2cm, below=0.7cm of agent, fill=green!8] (mcp) {
  \textbf{MCP / UCP / ACP Server} \texttt{(:8001)}\\[2pt]
  \texttt{/search\_products} $\cdot$ \texttt{/get\_product}\\
  \texttt{/ucp/...} $\cdot$ \texttt{/acp/...}
};

%% --- Data stores ---
\node[store, below left=0.8cm and 0.8cm of mcp, fill=yellow!12] (pg) {
  \textbf{PostgreSQL}\\21k products\\SQL filters\\hard constraints
};
\node[store, below=0.8cm of mcp, fill=purple!10] (neo) {
  \textbf{Neo4j KG}\\2.4k nodes\\8.5k edges\\17ms Cypher
};
\node[store, below right=0.8cm and 0.8cm of mcp, fill=red!8] (redis) {
  \textbf{Redis}\\Cache\\Session\\30ms hit
};

%% --- Arrows ---
\draw[arr] (agent) -- node[right, font=\scriptsize]{HTTP} (mcp);
\draw[arr] (mcp) -- (pg);
\draw[arr] (mcp) -- (neo);
\draw[arr] (mcp) -- (redis);
\end{tikzpicture}
\caption{End-to-end system architecture. The agent service (port~8000) handles
  conversation logic; the MCP server (port~8001) exposes protocol endpoints and
  routes all correctness-critical operations to three data stores.}
\label{fig:arch}
\end{figure}

\subsection{Protocol Layer}

The protocol layer exposes three endpoint families from a single backend:

\paragraph{MCP (Model Context Protocol).}
Four typed tool calls---\texttt{search\_products}, \texttt{get\_product},
\texttt{add\_to\_cart}, \texttt{checkout}---each return a standard envelope
with \texttt{trace.timings\_ms}, \texttt{cache\_hit}, and
\texttt{catalog\_version} fields, allowing agent developers to measure
latency sources, verify cache behavior, and detect catalog staleness
without backend access.

\paragraph{UCP (Universal Commerce Protocol).}
Endpoints \texttt{/ucp/search}, \texttt{/ucp/get-product},
\texttt{/ucp/get\_cart}, and \texttt{/ucp/checkout} follow Google's UCP
schema~\cite{UCPGoogle2026}, supporting shipping option selection by US state
and multi-item cart management.

\paragraph{ACP (Agent Commerce Protocol).}
Product feed, checkout session lifecycle, and webhook endpoints follow
OpenAI's ACP schema~\cite{ACPOpenAI2025}.
A single environment variable (\texttt{COMMERCE\_PROTOCOL}) selects between
UCP and ACP at deploy time, and tool schemas are exposed in OpenAI, Gemini,
and Claude function-calling formats, allowing any LLM agent with function-calling
support to transact directly.

\subsection{Data Layer}

Three data stores serve distinct roles:

\textbf{Supabase PostgreSQL} stores the authoritative product catalog
(21,000+ items; 1,490+ laptops with 12 structured spec fields per product:
CPU, RAM, GPU, storage type, display size, battery life, weight, OS, price,
brand, review count, and rating).
All correctness-critical operations---constraint filtering, brand exclusion,
progressive relaxation---execute as SQL queries against this store.

\textbf{Neo4j Knowledge Graph} encodes product relationships across
2,400 nodes and 8,500 edges in four relationship types:
\textsc{SIMILAR\_TO} (same use-case substitutes at similar price),
\textsc{COMPATIBLE\_WITH} (accessory compatibility),
\textsc{MANUFACTURED\_BY} (brand graph), and
\textsc{USE\_CASE} (product-to-use-case mapping).
The KG is used for ``see similar items'' (SIMILAR\_TO traversal) and
``upgrade suggestions'' (BETTER\_THAN traversal) without LLM inference.

\textbf{Upstash Redis} caches search results (5-minute TTL), product
summaries (5 minutes), price data (60 seconds), and LLM-generated
narratives (10 minutes), achieving $\sim$12$\times$ latency reduction
on cache hits (36~ms vs.\ 446~ms cold read, measured over $n = 10$ runs).

%% --------------------------------------------------------------------------
%% Algorithmic subsections — part of The Merchant Agent
%% --------------------------------------------------------------------------

\subsection{Query Rewriting}
\label{sec:rewriter}

The query rewriter (\texttt{agent/query\_rewriter.py}) transforms raw user
messages into annotated strings before any LLM call.
Its primary role is to inject structured hints---enclosed in \texttt{[\,]}---that
the downstream extraction LLM can interpret reliably.

Key transformations include:

\begin{itemize}
\item \emph{Common-sense enrichment:} ``for my son'' $\rightarrow$
  \texttt{[use\_case: school]}; ``Final Cut Pro'' $\rightarrow$
  \texttt{[brand: Apple] [use\_case: video\_editing]}.

\item \emph{Contradiction detection:} ``Dell Mac'' $\rightarrow$
  \texttt{[note: contradictory requirements---macOS is Apple-only]};
  ``gaming Chromebook'' $\rightarrow$ annotation noting ChromeOS incompatibility.

\item \emph{User expertise detection:} Non-technical signals
  (``I just need it for Netflix'', ``doesn't freeze'') without spec vocabulary
  inject \texttt{[note: lifestyle/non-technical user]}, causing the response
  generator to omit abbreviations.

\item \emph{Typo interception:} ``1TB RAM'' triggers a clarifying question
  before reaching the LLM (consumer laptops top out at 64--128~GB RAM).
\end{itemize}

These annotations are auditable plain text; the rewriter introduces no
opaque intermediate representations.

\subsection{Criteria Extraction}
\label{sec:extraction}

Criteria extraction maps $m'$ to a set of slot-value pairs
$\mathcal{S} = \{(s_i, v_i)\}$ where $s_i$ is a slot name from the domain
schema (e.g., \texttt{budget}, \texttt{min\_ram\_gb}, \texttt{excluded\_brands})
and $v_i$ is the extracted value.

The system uses two extraction paths:

\paragraph{Primary: LLM extraction.}
A structured prompt (\texttt{CRITERIA\_EXTRACTION\_PROMPT}) lists all valid
slot names and value types for the active domain.
Returned slot names are normalized through \texttt{\_SLOT\_NAME\_ALIASES}
(correcting LLM-returned variants like ``ram'' $\rightarrow$ \texttt{min\_ram\_gb})
and brand values through \texttt{\_BRAND\_VALUE\_ALIASES}
(``mac'' / ``macbook'' $\rightarrow$ ``Apple'').

\paragraph{Fallback: Regex extraction.}
A deterministic regex pass handles high-frequency patterns:
price ranges (\texttt{\$[0-9]+}), brand names, screen sizes, and RAM values.
The fallback fires when the LLM path returns no slots or the API is unavailable.

\paragraph{Brand exclusion.}
Indirect exclusion phrases (``bad experience with HP'', ``steer clear of ASUS'')
are processed by a dedicated micro-LLM call
(\texttt{\_extract\_excluded\_brands\_semantic}) that specifically classifies
negation intent.
This handles the 40+ indirect exclusion phrasings that a regex cannot cover.

\subsection{Hard Constraint Filtering (SQL)}

Extracted slot values map to a SQL WHERE clause (Algorithm~\ref{alg:sql}).
Brand exclusions are enforced both in SQL (\texttt{brand NOT ILIKE}) and
in a Python post-filter that checks the product title---necessary for
products where the database \texttt{brand} field stores condition
(``New'', ``Recertified'') rather than manufacturer name.

\begin{algorithm}[t]
\caption{SQL Hard Constraint Filtering}
\label{alg:sql}
\begin{algorithmic}[1]
\State \textbf{Input:} slot-value pairs $\mathcal{S}$, product catalog
\State Build WHERE clause from $\mathcal{S}$:
  \texttt{price} $\leq$ budget,
  \texttt{brand NOT ILIKE} excluded,
  \texttt{ram\_gb} $\geq$ min\_ram
\State Execute: \texttt{SELECT * FROM products WHERE ... ORDER BY rating DESC}
\State Post-filter: re-check brand exclusion against product title
\State \Return top-$k$ results (with progressive relaxation if $k < 3$)
\end{algorithmic}
\end{algorithm}

\subsection{Soft Constraint Ranking (Embeddings)}

For use-case-driven queries (\textit{gaming}, \textit{machine learning},
\textit{video editing}), products passing hard filters are ranked by cosine
similarity between the query embedding and pre-computed product embeddings.
Boolean flags (\texttt{good\_for\_gaming}, \texttt{good\_for\_ml},
\texttt{good\_for\_creative}) accelerate this path for common use cases,
avoiding embedding computation at query time.

\subsection{Progressive Relaxation}

When hard constraints produce fewer than three results, the system
applies three progressive relaxation steps while preserving
correctness-critical filters (price ceiling, brand exclusions):

\begin{enumerate}
\item \textit{First}, drop the \texttt{min\_ram\_gb} constraint.
\item \textit{Second}, drop the \texttt{storage\_type} constraint.
\item \textit{Third}, drop the \texttt{min\_screen\_size} constraint.
\end{enumerate}

The user always sees real, purchasable products; the system never
returns an empty result set or fabricated alternatives.

\subsection{Knowledge Graph Query Resolution}
\label{sec:kg}

Complex intent queries---``show me similar items,'' ``find a better option,''
``what accessories are compatible with this?''---are resolved via Cypher
traversal rather than LLM inference (Algorithm~\ref{alg:kg}).
This returns in $\sim$17~ms and requires no LLM call.
Full product data for the returned IDs is fetched from PostgreSQL.
The KG path gracefully degrades to SQL-based search when Neo4j is
unavailable (\texttt{is\_available()} returns false), ensuring no
hard dependency on the graph store.

\begin{algorithm}[t]
\caption{Knowledge Graph Traversal (SIMILAR\_TO)}
\label{alg:kg}
\begin{algorithmic}[1]
\State \textbf{Input:} product\_id $pid$, hop count $h = 1$
\State Query: \texttt{MATCH (p:Product \{product\_id: \$pid\})}
  \texttt{-[r:SIMILAR\_TO]->(q:Product)}
\State \Return $q$.\textit{product\_id} ordered by $r$.\textit{score} DESC, LIMIT~6
\State Fetch full product records for returned IDs from PostgreSQL
\State \textit{Fallback:} if Neo4j unavailable, use SQL embedding similarity
\end{algorithmic}
\end{algorithm}

\subsection{Best-Value Scoring}

The best-value pick uses a weighted scoring function tuned to the
detected use case:

\begin{equation}
\text{score}(p) = w_{\text{price}} \cdot r_p +
                  w_{\text{rating}} \cdot \hat{r} +
                  w_{\text{vol}} \cdot \log(n+1) +
                  w_{\text{spec}} \cdot \phi(p)
\end{equation}

where $r_p$ is the normalized price score (lower price $\rightarrow$ higher score),
$\hat{r}$ is the normalized star rating, $n$ is the review count, and
$\phi(p)$ is a use-case-specific spec score (GPU weight for gaming,
battery weight for travel, price weight for students).

Use-case weights are defined in \texttt{\_USE\_CASE\_WEIGHTS}:
gaming and ML assign 60\% weight to $\phi(p)$ and 15\% to price;
student queries assign 45\% to price and 20\% to specs.

\subsection{Intent Taxonomy and Coverage}

We identify thirteen primary intent classes, detected through a layered
routing system.

\begin{table}[t]
  \caption{Intent taxonomy and detection mechanism. Fast-path intents
  are detected by keyword matching without an LLM call.}
  \label{tab:intents}
  \begin{tabular}{lll}
    \toprule
    Intent & Example phrasing & Detection \\
    \midrule
    Search/Find & ``find a gaming laptop under \$1k'' & Domain + extraction \\
    Recommend & ``just pick one for me'' & Slot: \texttt{wants\_recs} \\
    Compare & ``X vs Y'', ``differences between'' & Fast-path keywords \\
    Refine & ``show me something cheaper'' & Fast-path keywords \\
    See Similar & ``show me similar items'' & KG SIMILAR\_TO \\
    Add to Cart & ``add the second one'' & Regex + ordinal parse \\
    Research & ``explain features'' & Keyword match \\
    Checkout & ``checkout'', ``pay'' & Keyword match \\
    Pros/Cons & ``tell me more about these'' & Fast-path keywords \\
    Targeted Q\&A & ``which has best battery?'' & Fast-path keywords \\
    Rate & ``5 stars'' & Exact match \\
    New Search & ``start over'' & LLM intent router \\
    Domain Switch & ``show me books instead'' & LLM refinement \\
    \bottomrule
  \end{tabular}
\end{table}

The fast-path router (Table~\ref{tab:intents}) intercepts known intent patterns
before any LLM call.
For the ``compare'' intent, the keyword set was extended to cover natural
paraphrases: ``differences between,'' ``contrast,'' ``side by side,''
``how do they differ,'' and ``what sets them apart,'' addressing a gap
identified in user testing where comparison-intent queries were misclassified
as new searches.

The ``add to cart'' intent handler parses ordinal references
(``first,'' ``second,'' ``third'' $\rightarrow$ indices 0, 1, 2) against
\texttt{last\_recommendation\_data}, falls back to the frontend context
tag (\texttt{[ctx:id1,id2,...]}) when no ordinal is present, and calls
\texttt{session\_manager.add\_favorite()} to update the cart.

\subsection{Protocol Unification: Design and Implications}
\label{sec:protocol-unification}

Implementing MCP, UCP, and ACP on a single backend was motivated by the
protocol fragmentation challenge described in Section~\ref{sec:introduction}.
Three design principles proved necessary for clean unification:

\paragraph{Schema normalization.}
Each protocol uses different field names for identical concepts (e.g., UCP
\texttt{product\_id} vs.\ ACP \texttt{item\_id}).
A canonical internal schema is defined with per-protocol adapters at
the endpoint layer.

\paragraph{Lifecycle compatibility.}
ACP requires a checkout session lifecycle with webhooks; UCP does not.
ACP's session semantics are implemented as an optional overlay---merchants
that don't need webhook callbacks simply ignore them.

\paragraph{Observable envelopes.}
All three protocols expose \texttt{trace.timings\_ms}, \texttt{cache\_hit},
and \texttt{catalog\_version} in every response.
This allows agent developers to measure end-to-end latency, diagnose
cache behavior, and detect catalog staleness programmatically---capabilities
important for production agent deployment but absent from the protocol
specifications.

\subsubsection{Implication for Merchants}

The implementation shows that per-protocol implementation cost is
lower than typically assumed.
Once a canonical product schema and search layer exist, adding a new
protocol endpoint primarily requires: (1) a request adapter (field renaming),
(2) a response formatter (protocol-specific envelope), and (3) authentication
middleware.
The heavy lifting---product retrieval, filtering, caching, KG traversal---is
shared.

This suggests merchants should prioritize building a well-structured internal
product API first, then surface it through multiple protocol adapters,
rather than attempting to natively implement each protocol's full semantics.

%% ==========================================================================
\section{Evaluation}
%% ==========================================================================

\subsection{Benchmark Design and Ground Truth Construction}

The evaluation requires measuring two distinct properties:
(1) \emph{constraint correctness}---did the system enforce the user's hard
requirements (price, brand exclusions, product type)?---and
(2) \emph{response quality}---is the generated recommendation helpful,
accurate, and appropriately explained?

These properties require different evaluation methods, and conflating them
is a fundamental flaw in evaluations that use an LLM judge for both.

\paragraph{Deterministic ground truth (hard constraints).}
For each of the 43 benchmark queries, expected filter values are specified
(\texttt{price\_max\_cents}, \texttt{min\_ram\_gb}) and expected exclusions
(\texttt{excluded\_brands}) derived by executing the constraints directly
against the PostgreSQL database:

\begin{verbatim}
SELECT id FROM products
WHERE product_type = 'laptop'
  AND price <= 1000
  AND brand NOT ILIKE '%HP%'
  AND good_for_gaming = true
ORDER BY rating DESC, review_count DESC
LIMIT 10;
\end{verbatim}

This ground truth is independent of any LLM---violations are verifiable
objective failures.
This approach eliminates the circularity problem of using one LLM to generate
reference answers and another to judge whether the system matches them.

\paragraph{Embedding-based soft constraint ground truth.}
For use-case queries (``best for machine learning'', ``good for video editing''),
expected top-$k$ products are computed offline as those with highest cosine
similarity to the query embedding in the pre-computed product embedding space,
stored as reference sets.

\paragraph{LLM-as-judge quality scoring.}
For response narrative quality, G-Eval~\cite{Liu2023GEval} is applied with a
custom rubric evaluating engagement (0--4), tone calibration (0--3), and
factual accuracy/coherence (0--3), normalized to $[0,1]$.
The judge is GPT-4o-mini at temperature 0 for reproducibility.
It evaluates whether explanations are factually grounded in the returned
products---not whether they resemble a GPT-generated reference.
Calibration examples anchor scores at 0.0, 0.5, and 1.0 to reduce judge variance.
All 43 queries execute concurrently (semaphore=6) for throughput.

\paragraph{Composite scoring formula.}
The final per-query score combines deterministic and quality components:
\begin{equation}
  s = 0.40\,s_{\text{type}} + 0.20\,s_{\text{brand}} + 0.10\,s_{\text{filter}}
    + 0.05\,s_{\text{stock}} + 0.25\,s_{\text{quality}}
  \label{eq:geval}
\end{equation}
Weights for components that are not applicable to a given query (e.g., brand
exclusion when no brand constraint exists) are redistributed to
$s_{\text{quality}}$, raising its effective weight to 55--60\%.
This redistribution means text-only baselines (GPT, Gemini) are evaluated
only on type and quality---they cannot be penalized for brand or filter
violations they are structurally unable to enforce.
The 43 queries split into 28 \emph{specified} (user provides $\geq 1$ hard
constraint) and 15 \emph{underspecified} (elicitation required).
Pass threshold: score $\geq 0.5$.

\subsection{Multi-Turn Evaluation Design}
\label{sec:multiturn-eval}

Single-turn evaluations cannot assess the properties that distinguish agentic
systems from stateless LLM calls: constraint persistence, preference updates,
and brand-exclusion reversals across turns.
A 10-scenario multi-turn benchmark is designed covering the following interaction
patterns: (1)~constraint accumulation over 4 turns; (2)~use-case pivot
(gaming $\rightarrow$ video editing); (3)~brand exclusion persistence across all
subsequent turns; (4)~vague-to-specific elicitation (one-word start);
(5)~comparison follow-up in context; (6)~downward price refinement;
(7)~budget increase overwrite (update, not accumulate); (8)~brand exclusion
reversal (``HP is fine now''); (9)~full 5-turn session simulating a
medical-student purchase flow; and (10)~dense multi-constraint first turn
followed by a clarification. Each scenario contains 3--5 scripted turns.

\paragraph{Scoring.}
The multi-turn score combines a judge score with deterministic constraint checks:
\begin{equation}
  s_{\text{MT}} = 0.55\,s_{\text{constraint}} + 0.45\,s_{\text{judge}}
  \label{eq:multiturn}
\end{equation}
$s_{\text{constraint}}$ is computed from deterministic checks against the
system's final-turn product list: brand exclusions present/absent, budget
ceiling respected.
$s_{\text{judge}}$ is GPT-4o-mini (temperature 0) evaluating the full
conversation transcript on three criteria: constraint satisfaction (0--4),
preference tracking accuracy (0--3), and response appropriateness (0--3).
A critical fairness rule prevents penalizing \emph{silent} enforcement:
the judge checks only whether final products comply with constraints, not
whether the system verbalized constraint acknowledgment.
Pass threshold: $\geq 0.5$.

\paragraph{Fair catalog-bound baselines.}
Plain GPT and Gemini baselines cannot be fairly compared to IDSS because
they lack database access: without a product DB, any text-only baseline is
penalized for an architectural gap rather than a quality gap.
To correct for this, before each baseline turn we call IDSS's search
endpoint (\texttt{/chat?k=0}) to retrieve the top-10 products for the
current query state, then inject those products into the baseline's system
prompt as ``CATALOG---recommend ONLY from these products.''
All systems therefore recommend from an identical, live catalog snapshot,
isolating the architecture comparison from database differences.

\subsection{Latency Measurement}

The full request lifecycle is instrumented with phase-level timing logged to
\texttt{trace.timings\_ms} in each MCP response envelope.
Table~\ref{tab:latency} shows representative latencies for each phase.

\begin{table}[t]
  \caption{Measured latencies by pipeline phase ($n = 10$ runs per phase,
  direct-call benchmark). LLM calls dominate total latency; KG traversal
  and Redis cache hits are negligible.}
  \label{tab:latency}
  \begin{tabular}{lrr}
    \toprule
    Phase & Median (ms) & Notes \\
    \midrule
    CPU phases (rewrite, filter)  &   $<$1  & Pure Python, no I/O \\
    Domain detection — fast path  &   $<$1  & Keyword dict lookup \\
    Domain detection — LLM        & $\sim$1{,}630 & gpt-4o-mini; fires only for ambiguous queries \\
    Criteria extraction (LLM)     & 1{,}927 & Primary bottleneck \\
    Question generation (LLM)     & 1{,}645 & Preference elicitation \\
    Post-rec intent detection      &   556   & gpt-4o-mini classification \\
    Rec.\ narrative (LLM)         & 2{,}256 & Recommendation explanation \\
    Comparison narrative (LLM)    & 5{,}714 & Longest single phase \\
    Filter refinement (LLM)       & 2{,}187 & gpt-4o-mini RefinementClassification \\
    SQL search (cache miss)       &   446   & Supabase PostgreSQL round-trip \\
    SQL search (cache hit)        &    36   & Redis GET + deserialise \\
    KG re-ranking overlay         &   207   & FAISS + graph candidates + scoring \\
    Total (cache miss, rec.)      & $\sim$4{,}600 & Criteria + SQL + narrative \\
    Total (cache hit, rec.)       & $\sim$2{,}300 & Criteria + Redis \\
    \bottomrule
  \end{tabular}
\end{table}

LLM inference accounts for 85--92\% of total latency on cache misses.
The KG re-ranking overlay (207~ms median) is 11--28$\times$ faster than
equivalent LLM-based narrative generation (2,256--5,714~ms median) and produces
deterministic, auditable results---a strong argument for routing structural
queries (substitutes, best value) through the KG rather than prompting.

Redis caching reduces effective latency $\sim$12$\times$ on repeated queries
(446~ms cold vs.\ 36~ms cached; popular product categories, frequently compared
spec configurations).

\subsection{Single-Turn Quality Results}

Table~\ref{tab:results} compares IDSS against three baselines on the 43-query
shared benchmark.
GPT+Catalog and Gemini+Catalog receive IDSS's catalog per query (catalog
grounding = 1.0); Baseline~B uses its own search pipeline.
Brand and filter scores are reported only for IDSS and Baseline~B---text-only
baselines return free text and cannot be evaluated on these dimensions
(marking them N/A, not zero).

\begin{table}[t]
  \caption{Single-turn G-Eval results ($n = 43$ shared queries, pass
  threshold $\geq 0.5$, SE $= \hat{\sigma}/\sqrt{n}$; IDSS re-evaluated April 2026,
  baselines from March 2026 run; single run---multi-seed runs with $\pm$ SD
  planned as part of the expanded 225-query benchmark).
  Brand/Filter: deterministic compliance scores;
  N/A for text-only systems (free text, not penalized).
  Quality: LLM-judge narrative score (25\% weight base, up to 60\%
  when brand/filter components are N/A).}
  \label{tab:results}
  \begin{tabular}{lllllll}
    \toprule
    System & Avg $\pm$ SE $\uparrow$ & Pass\% $\uparrow$ & Brand $\uparrow$ & Filter $\uparrow$ & Quality $\uparrow$ \\
    \midrule
    IDSS (ours)    & $0.683 \pm 0.020$ & 90.7\% & 1.000 & 1.000 & 0.465 \\
    GPT+Catalog    & $0.662 \pm 0.032$ & 65.1\% & N/A   & N/A   & 0.381 \\
    Gemini+Catalog & $0.671 \pm 0.030$ & 58.1\% & N/A   & N/A   & 0.370 \\
    Baseline~B     & $0.271 \pm 0.033$ & 11.6\% & 0.000 & 0.000 & 0.088 \\
    \bottomrule
  \end{tabular}
\end{table}

The current IDSS results reflect three targeted improvements over a v8 baseline
(avg.\ 0.647, pass rate 83.7\%): (1)~a RAM-disclosure bug that added false
``no 16\,GB RAM available'' notes when RAM data was absent from the response,
misleading the judge into lower accuracy scores; (2)~a recommendation-explanation
refactor that now passes only structured user preferences---not raw chat
messages---to the narrative LLM, eliminating confusion caused by typos and
all-caps user input; and (3)~query-rewriter non-technical user detection,
which suppresses spec abbreviations for lifestyle-oriented queries.
These three fixes, along with enhanced negation handling (``steer clear of HP'',
``bad experience with Dell''), account for the improvement to 90.7\% pass rate.
The gain over v8 (+0.036 avg score, +6.7~pp pass rate) is approximately
1--2 combined standard errors and is practically meaningful given $n = 43$.

IDSS leads all fair baselines on pass rate (+25.6\,pp vs.\ GPT+Catalog,
+32.6\,pp vs.\ Gemini+Catalog, +79.1\,pp vs.\ Baseline~B) and quality score
($+0.084$ and $+0.095$ above catalog-aware GPT and Gemini, respectively).
The aggregate score gap between IDSS (0.683) and GPT+Catalog (0.662) is
0.021 points---within one standard error of either system
(SE$_\text{IDSS}$\,=\,0.020, SE$_\text{GPT}$\,=\,0.032)---because
text-only systems are evaluated only on type and quality: they cannot lose
points for brand or filter violations they are structurally unable to enforce.
The meaningful comparison is therefore not the aggregate score but the
quality sub-score (0.465 vs.\ 0.381, $+0.084$) and the brand/filter
compliance that text-only systems cannot provide at all.
IDSS's brand score of 1.000 and filter score of 1.000 reflect SQL-level
constraint enforcement; Baseline~B's 0.000 on both reveals that its pipeline
does not reliably apply these filters.

The system still fails on five query categories:
(1) products genuinely absent from the catalog (ultra-high-spec configs);
(2) marketplace risk questions requiring external information;
(3) iPad/Surface pivot requests (catalog contains only laptops);
(4) split budget queries (monitor + laptop under a single budget);
(5) DaVinci Resolve with gaming-avoidance (search returns accessories).
These are catalog-level or scope limitations that cannot be resolved at the
agent layer.

\subsection{Multi-Turn Constraint Results}
\label{sec:multiturn-results}

Table~\ref{tab:multiturn} reports multi-turn evaluation results across 10 scripted
scenarios.
IDSS leads all fair baselines with an average score of $0.604 \pm 0.013$, edging
catalog-bound GPT ($0.595 \pm 0.015$) and outperforming Baseline~B
($0.554 \pm 0.005$) and Gemini ($0.532 \pm 0.018$, separate run).
The gap between IDSS and catalog-bound GPT ($0.009$) falls within one combined
standard error and is not statistically significant; the margins versus
Baseline~B ($0.050$, ${\approx}3.6$ combined SEs) and Gemini ($0.072$,
${\approx}3.3$ combined SEs) are statistically significant.

\begin{table}[t]
  \caption{Multi-turn evaluation results (10 scenarios, 3--5 turns each;
  single run---multi-seed runs with $\pm$ SD planned as part of expanded benchmark).
  Scoring: $0.55 \times$ constraint (deterministic) $+ 0.45 \times$ judge
  (GPT-4o-mini, temperature~0). GPT and Gemini are catalog-bound
  (same product pool as IDSS per turn). Gemini results from a separate run.
  Pass threshold: $\geq 0.5$.}
  \label{tab:multiturn}
  \begin{tabular}{lrr}
    \toprule
    System & Avg $\pm$ SE $\uparrow$ & Pass/10 $\uparrow$ \\
    \midrule
    IDSS (ours)    & $0.604 \pm 0.013$ & 10/10 \\
    GPT+Catalog    & $0.595 \pm 0.015$ & 10/10 \\
    Baseline~B     & $0.554 \pm 0.005$ & 10/10 \\
    Gemini+Catalog & $0.532 \pm 0.018$ &  9/10 \\
    \bottomrule
  \end{tabular}
\end{table}

The 10 scenarios span six constraint-persistence categories:
(1)~constraint accumulation across turns (brands, RAM, budget added incrementally);
(2)~mind-change reversal (user first excludes then re-allows a brand);
(3)~brand-exclusion persistence (``no HP'' must survive follow-up clarifications);
(4)~vague-to-specific refinement (user narrows from category to exact spec);
(5)~comparison follow-up (user requests side-by-side after initial recommendations);
(6)~budget overwrite (new price ceiling replaces old, not added on top).

IDSS's architecture enforces constraints deterministically:
brand exclusions are SQL-level predicates persisted in Redis across turns, so they
cannot be silently dropped by LLM inference.
Baseline~B achieves low variance ($\pm 0.005$) because its responses
consistently match a low-quality baseline across all scenarios.
Gemini fails Scenario~8 (exclusion reversal: ``actually HP is fine'')
because its stateless context management does not detect the un-exclusion signal.

\subsection{Trust \& Safety Evaluation}
\label{sec:trust}

Standard quality metrics measure \emph{what} a system recommends; they do not
measure whether it behaves safely under adversarial or underspecified inputs.
Three Trust \& Safety pillars are evaluated, each measured deterministically.
Together, they assess whether the system provides the \emph{guarantees}
central to the design: enforcing constraints when possible, persisting them
across turns, and disclosing honestly when they cannot be met.

\paragraph{Inventory integrity (catalog grounding).}
Every product ID returned by the system is looked up in PostgreSQL and verified
to exist at the stated price.
Score = fraction of recommended products that exist in the live catalog.
IDSS, GPT+Catalog, and Gemini+Catalog all score 1.000 because all three use
IDSS's PostgreSQL-backed product layer as their source.
A stateless GPT call (no DB access) would score $\approx$0.00 on this metric.

\paragraph{Preference fidelity (constraint drift rate).}
The 10 multi-turn scenarios are annotated with \texttt{per\_turn\_constraints}
specifying which brand exclusions and budget ceilings are active at each turn.
After running each scenario end-to-end, every turn where a constraint was active
\emph{and} products were returned is checked:
\[
  \text{drift\_rate} = \frac{\text{violations}}{\text{applicable (turn, constraint) pairs}}
\]
This measures whether constraints \emph{persist} across turns, not merely whether
they are satisfied in the final turn---a stronger property than final-turn compliance.
All three systems score 0.000 in the primary run; a secondary run revealed
GPT+Catalog drift = 0.042 on Scenario~7 (budget-increase overwrite), consistent
with GPT relying on prompt context versus IDSS's Redis-persisted typed slot dict.

\paragraph{Honest scoping (impossible query disclosure).}
Eight \emph{catalog-impossible} queries are constructed---requests whose stated
constraints cannot be satisfied by any product in the catalog
(e.g., ``gaming laptop with RTX~4090, budget \$150'').
Each query carries a \texttt{hard\_budget\_usd} field enabling deterministic scoring:
\begin{itemize}
\item \textbf{1.0 (Pass):} Response explicitly discloses the constraint gap
  (``I couldn't find...'', ``I relaxed your budget because...'').
\item \textbf{0.5 (Soft fail):} Products returned within budget but no disclosure---silent relaxation.
\item \textbf{0.0 (Hard fail):} Products returned, at least one exceeding the stated budget by ${>}5\%$.
\end{itemize}
The group pass threshold is raised to 0.7 (versus the global 0.5), so only explicit
disclosure earns a pass.
IDSS scores 0.925: its progressive relaxation pipeline generates explicit
disclosure notes (``I relaxed your budget constraint because no laptops matched
exactly'') that trigger the 1.0 tier on six of eight queries.
GPT+Catalog and Gemini+Catalog both score below 0.6: they receive IDSS's
already-relaxed product list but have no mechanism to detect that relaxation
occurred, so they recommend without disclosure.

\begin{table}[t]
  \caption{Trust \& Safety evaluation across three pillars ($n = 8$ impossible
  queries for disclosure; $n = 10$ scenarios for drift).
  $\uparrow$~higher is better; $\downarrow$~lower is better.
  N/A: metric not applicable to this system's architecture.}
  \label{tab:trust}
  \begin{tabular}{lrrr}
    \toprule
    System & Grounding $\uparrow$ & Disclosure $\uparrow$ & Drift $\downarrow$ \\
    \midrule
    IDSS (ours)    & 1.000 & \textbf{0.925} & \textbf{0.000} \\
    GPT+Catalog    & 1.000 & 0.588          & 0.000          \\
    Gemini+Catalog & 1.000 & 0.550          & 0.000          \\
    Baseline~B     & N/A   & N/A            & N/A            \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:trust} summarizes all three pillars.
The disclosure gap ($+0.338$ vs.\ GPT+Catalog, $+0.375$ vs.\ Gemini+Catalog)
is the clearest architectural differentiator in this evaluation: IDSS detects
and discloses catalog constraint gaps at the system level; catalog-bound
baselines cannot, because they are unaware of the relaxation that produced
the products they received.

\subsection{Agentic System vs.\ Plain LLM Baseline}

Tables~\ref{tab:results} and~\ref{tab:multiturn} quantify IDSS's advantage on
single-turn and multi-turn benchmarks.
Table~\ref{tab:vs-gpt} summarizes the qualitative capability differences
between the agentic architecture and a plain GPT-4o-mini call with the same
user message.

\begin{table*}[t]
  \caption{Capability comparison: IDSS agentic system vs.\ plain GPT-4o-mini call.}
  \label{tab:vs-gpt}
  \begin{tabular}{lll}
    \toprule
    Capability & Plain GPT & IDSS \\
    \midrule
    Brand exclusion enforcement  & Unreliable (LLM ignores $\sim$15\% of exclusions) & Deterministic (SQL NOT ILIKE) \\
    Product spec accuracy        & Hallucinated  & Database-sourced, auditable \\
    Cross-turn preference memory & None (stateless) & Redis-persisted session state \\
    Progressive relaxation       & None          & 3-step relaxation, always returns results \\
    Entropy-based question selection & None      & Slot-information-gain maximization \\
    Query response latency (cached) & 500ms+    & $\sim$36~ms (Redis hit) \\
    Complex query resolution     & LLM inference (2,694--6,832~ms, hallucination risk) & KG traversal (17~ms, deterministic) \\
    Catalog grounding            & 0\% (no DB access) & 100\% (DB-sourced per query) \\
    Multi-turn constraint pass rate & Not evaluated & 10/10 (Table~\ref{tab:multiturn}) \\
    Evaluable against ground truth & No (text only) & Yes (filter dicts, product IDs, session state) \\
    \bottomrule
  \end{tabular}
\end{table*}

The critical distinction is architectural: IDSS uses LLMs as one component in a
larger system, delegating correctness-critical operations (filtering, deduplication,
session persistence) to deterministic components.
A plain LLM call collapses all of these into a single inference step, trading
reliability and verifiability for simplicity.
For consumer-facing product recommendation, where an incorrect brand exclusion
or hallucinated price can directly harm the user, this reliability gap is significant.
When GPT-4o-mini receives the same catalog per query (GPT+Catalog in
Table~\ref{tab:results}), the aggregate score gap narrows to $0.021$ points,
but IDSS retains a $+0.084$ quality advantage ($0.465$ vs.\ $0.381$)
because it selects products that better match the user's stated constraints.

%% ==========================================================================
%% ==========================================================================
\section{Discussion}
%% ==========================================================================

\subsection{KG Utility and Limitations}

The knowledge graph delivers the clearest benefit on structural queries
(``show me similar items,'' ``find compatible accessories'') that have
deterministic answers given the product data. For queries requiring external
judgment (``is this price fair?'', ``will I be happy with this for five years?''),
the KG adds no value and those queries remain in the LLM path.

KG utility is directly gated on graph construction quality.
Edges must reflect real product relationships---compatibility specifications,
co-purchase data, or verified spec similarity---rather than heuristic rules.
In the current deployment, \textsc{SIMILAR\_TO} edges are computed from
embedding cosine similarity above a threshold, while \textsc{COMPATIBLE\_WITH}
edges are manually curated from manufacturer compatibility tables.
Improving automated edge construction from catalog metadata is a natural
direction for future work.

\subsection{Interview vs.\ Zero-Shot Elicitation}

The interview-style preference elicitation adds 1--3 turns before
recommendations appear for underspecified queries.
Users with explicit, well-specified preferences bypass elicitation entirely:
the interview threshold (skip questions if $\geq$1 substantive criterion
is present) ensures that single-constraint queries such as ``gaming laptop
under \$1000'' proceed directly to search.
For genuinely underspecified queries, targeted elicitation narrows the
feasible product set before search runs, reducing the chance of irrelevant
recommendations.
Future work should explore dynamically adjusting this threshold based on
detected user expertise and session history.

\subsection{Protocol Standardization Outlook}

MCP, UCP, and ACP each address real deployment needs, but they diverge
on field names and lifecycle semantics in ways that create unnecessary
per-protocol engineering work.
The key takeaway from this implementation is that \emph{schema alignment}
(agreeing on canonical names for fields such as price, inventory, and
product type) would eliminate most of the adapter overhead.
The remaining differences---primarily lifecycle semantics for ACP webhooks
versus UCP's stateless model---are architectural choices that may converge
as deployment patterns mature.
Merchants are best served by building a well-structured internal product
API first and surfacing it through protocol adapters, rather than
natively implementing each protocol's full semantics.

%% ==========================================================================
\section{Conclusion}
%% ==========================================================================

IDSS is a merchant agent framework that unifies three competing
agentic commerce protocols (MCP, UCP, ACP) behind a single backend.
The central finding is not that the agentic architecture produces better
language---on aggregate score, IDSS is statistically tied with catalog-aware
GPT and Gemini (gap $< 1$ SE).
The finding is that it provides \emph{guarantees}: brand exclusions that
cannot be silently dropped, filter constraints that are SQL-verified rather
than inferred, and preference state that persists across turns.
These properties are architecturally impossible for a stateless LLM call,
regardless of prompt quality.

Five key empirical findings emerge:
(1) the KG re-ranking overlay (207~ms) resolves complex product queries 11--28$\times$ faster
than equivalent LLM narrative generation with deterministic, auditable results;
(2) on a 43-query single-turn benchmark, IDSS achieves 90.7\% pass rate
($0.683 \pm 0.020$), outperforming catalog-aware GPT-4o-mini ($0.662 \pm 0.032$)
on quality with brand and filter compliance of 1.000 versus N/A for text-only systems;
(3) on a 10-scenario multi-turn benchmark, IDSS ($0.604 \pm 0.013$)
leads catalog-bound GPT ($0.595$) and Baseline~B ($0.554$), with the
margin attributable to deterministic constraint persistence rather than
LLM output quality;
(4) Redis caching reduces median search latency $\sim$12$\times$ on repeated
queries (446~ms cold vs.\ 36~ms cached), making agent-mediated shopping
viable at production throughput;
(5) on a three-pillar Trust \& Safety benchmark (Table~\ref{tab:trust}),
IDSS achieves disclosure score 0.925 versus 0.588 (GPT+Catalog) and 0.550
(Gemini+Catalog), with zero constraint drift rate across all multi-turn
scenarios---a gap attributable to explicit relaxation disclosure and
SQL-level constraint enforcement that catalog-bound text systems cannot
replicate.

The emerging protocol fragmentation landscape (ACP, UCP, MCP) represents
both a risk and an opportunity for merchants.
Systems that abstract protocol differences behind a canonical data layer
can reach all major agent ecosystems without redundant engineering.
The full system and codebase are made available for replication
[URL omitted for blind review; will be provided in camera-ready version].

\begin{acks}
The authors thank their advisors and the host research laboratory for
advising and infrastructure support [details omitted for blind review].
AI writing assistance (large language model tools) was used in the
preparation of this manuscript, in accordance with the ACM RecSys 2026
Use of AI Policy.
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{idss-refs}

\appendix

\section{Intent Detection Keyword Sets}
\label{app:keywords}

The fast-path intent router uses the following keyword sets (evaluated in order,
short-circuit on first match):

\begin{itemize}
\item \textbf{Best value:} ``best value,'' ``get best,'' ``show me the best,'' ``best pick''
\item \textbf{Pros/Cons:} ``tell me more about these,'' ``pros and cons,'' ``worth the price,''
  ``trade-off,'' ``battery life on these''
\item \textbf{Targeted Q\&A:} ``which has the best,'' ``which is the most,'' ``which would you recommend,''
  ``which should I get,'' ``which should I pick''
\item \textbf{Compare:} ``vs,'' ``compare my,'' ``compare these,'' ``difference between,''
  ``differences between,'' ``how do they differ,'' ``contrast,'' ``side by side,''
  ``what sets them apart,'' ``detailed specs of each''
\item \textbf{See similar:} ``see similar,'' ``similar items,'' ``show me similar,''
  ``something similar,'' ``similar to''
\item \textbf{Refine:} ``refine my search,'' ``change my criteria,'' ``change budget,''
  ``different screen size,'' ``different brand''
\end{itemize}

Intents not matched by any fast-path keyword are routed to the LLM intent router
(\texttt{detect\_post\_rec\_intent}), which classifies the message as one of:
\texttt{compare}, \texttt{refine}, \texttt{new\_search}, \texttt{other}.

\section{G-Eval Rubric}
\label{app:geval}

Each response is scored 0.0--1.0 on five dimensions:

\begin{enumerate}
\item \textbf{Relevance (0--1):} Do the recommended products match the user's
  stated constraints (category, price range, use case)?
\item \textbf{Faithfulness (0--1):} Does the narrative accurately reflect the
  products' actual specifications (no fabricated specs or prices)?
\item \textbf{Helpfulness (0--1):} Does the response give the user actionable
  information to make a purchase decision?
\item \textbf{Tone (0--1):} Is the response appropriately warm for the user's
  apparent expertise level and emotional state?
\item \textbf{Constraint correctness (0--1):} Are hard constraints honored?
  (Excluded brands absent, price ceiling respected, correct product type returned.)
\end{enumerate}

The final score is the mean of the five dimensions.
Pass threshold is $\geq 0.5$.
Calibration examples anchor 0.0 (completely wrong category, wrong brand),
0.5 (correct category, one constraint violated), and 1.0 (all constraints
satisfied, accurate explanation).

\end{document}
